### 2022-10-03 elsman/manuscript.tex
submitted

% Created 2022-08-29 Mon 07:25
% Intended LaTeX compiler: pdflatex
\documentclass[preprint,numrefs,noinfo,sort&compress]{elsarticle}
  \usepackage[utf8]{inputenc}
\usepackage{url}
\usepackage[version=4]{mhchem}
\usepackage{chemmacros}[2016/05/02]
\usepackage{graphicx}
\usepackage{float}
\usepackage{color}
\usepackage{adjustbox}
\usepackage{amsmath}
\usepackage{siunitx}
\usepackage{textcomp}
\usepackage{wasysym}
\usepackage{latexsym}
\usepackage{amssymb}
\usepackage{lineno}
\usepackage{chemformula}
\usepackage{xr}
\usepackage{pifont}
\usepackage{longtable}
\usepackage[section]{placeins}
\usepackage{threeparttable}
\newcommand{\red}[1]{\textcolor{red}{#1}}
\chemsetup{formula = mhchem ,modules = {reactions,thermodynamics}}
\usepackage{xcolor}
\chemsetup[reactions]{tag-open= ( , tag-close = )}
\usepackage[capitalize]{cleveref}
\usepackage{minted}
\DeclareSIUnit\Td{Td}
\setminted[python]{frame=lines,fontsize=\scriptsize,xleftmargin=\parindent,linenos,breaklines}
\date{}
\title{}
\begin{document}

@@

\mainmatter
\section{Introduction}
\label{sec:org2a67457}
Zeolites are three dimensional crystalline structures containing tetrahedral Si or Al atoms connected by oxygen bridges. The International Zeolite Association (IZA) lists over 200 known zeolite frameworks that are described by dimensionality, pore shape and size, and Si/Al ratios \cite{baerlocher-database-nodate}. It is natural to characterize differences in zeolites by the size and shapes of the features present in the crystal. Rings are one common type of feature widely reported and used, both to characterize a zeolite crystal and as descriptors of the individual tetrahedral sites (T-sites) of a zeolite. Rings have been used to identify feasible zeolites topologies \cite{li-why-2014}, to describe the similarity and differences between zeolites \cite{curtis-statistical-2003,bermudez-calculation-2017}, to identify sites or voids of catalytic relevance \cite{li-first-principles-2018,kester-effects-2021}, and as machine learning finger prints of physicochemical properties \cite{helfrecht-new-2019}.

@@

Vertex symbols are a common way to describe the T-sites of a zeolite framework, and were first used to do so in 1997 by O'Keeffe and Hyde in 1997 \cite{okeeffe-vertex-1997}. The vertex symbol of a T-site contains the shortest rings associated with each of the six edges of the tetrahedron formed by the T-atom and its four bounded oxygen atoms. The ring sizes, and  their multiplicity, for opposite edges (edges of the tetrahedron that do not connect to the same oxygen atom) are grouped together. These grouped pairs are listed from smallest to largest forming the vertex symbol. \cref{fig:cha-vertex}(a) shows an example of the tetrahedron formed at the single symmetry distinct T-site in the CHA framework. The edges of the tetrahedron are labeled to aid in identification. \cref{fig:cha-vertex}(b-d) show the rings associated with each opposite pair of edges. The vertex symbol of T1 in CHA is thus able to be determined as 4\(\cdot\)4\(\cdot\)4\(\cdot\)8\(\cdot\)6\(\cdot\)8. For a T-site that contains a multiplicity of rings at one edge, that multiplicity would be represented as a subscript in the vertex symbol. An example would be the vertex symbol of 4\(\cdot\)6\(_{\text{2}} \cdot\)6\(\cdot\)6\(_{\text{3}} \cdot\)6\(_{\text{2}} \cdot\)6\(_{\text{3}}\) for T1 in AFI.

@@

\section{Methods}
\label{sec:org6431034}

\subsection{Finding Rings That do not Contain Shortcuts \label{section:goetzke}}
\label{sec:org42ddcba}

@@

\subsection{Finding Vertex Symbol Rings \label{section:vertex}}
\label{sec:org20e1097}

@@

\subsection{Finding Shortest Path Rings}
\label{sec:org42016a9}

@@

\subsection{Modified Shortcut Rings}
\label{sec:orgd5be89b}

@@

\subsection{Ordered Vertex Symbols \label{section:ov}}
\label{sec:org8f45030}

@@

\subsection{Determining All Ring Sizes Contained in a Zeolite Framework}
\label{sec:org79d764e}

@@

\section{Results}
\label{sec:orgcd96db0}
\subsection{Characterizing Rings in a Zeolite Graph}
\label{sec:org15e507d}
The IZA Database \cite{baerlocher-database-nodate} is a common reference used to identify all the rings in a zeolite framework, however it only lists the rings that define a channel (ex: 12-MR in AFI), or rings associated with the symbol of a T-site. These rings listed by the IZA are referred to as tabulated rings in the literature \cite{curtis-statistical-2003}. In some frameworks, other rings (cycles not containing shortcuts) exist that are not included in the list of tabulated rings. These 'untabulated' rings may still provide important topological information about a zeolite framework, or the local void environment around a T-site. \cref{fig:fw-counts} shows counts of frameworks containing each size ring from 3- to 18-MR using the Goetzke algorithm and the listed rings on the IZA database \cite{baerlocher-database-nodate}. There are slight differences in the counts up to 6-MRs, but the main divergence takes place as we get to ring sizes \textgreater6-MR.

@@

\subsection{Characterizing Frameworks by Rings}
\label{sec:orgb71506d}
With the addition of our modified shortcut definition, we have four ring finding conventions to compare, as well as including the tabulated rings from the IZA Database. \cref{fig:ring-counts} shows how many frameworks contain each size ring found using the various ring counting conventions from 3- to 18-MRs. This plot highlights the differences in the conventions and shows that a topological description of a framework based on rings will depend on the way that you define a ring. In general, a hierarchy of ring sizes found by each convention is: all rings not containing a shortcut \textgreater this work \textgreater shortest path rings \textgreater vertex symbol rings. While the IZA listed rings includes the vertex symbol rings, and a selection of general rings \cite{baerlocher-database-nodate}.

@@

One drawback to using a ring convention based on connectivity and shortcuts is the exclusion of cycles that don't fit this definition, but still exhibit properties like rings. This is a trade-off between well-defined connectivity rules, and the inclusion of particular void environments that may still have important applications. These shortcut containing cycles can display chemical and/or geometric properties consistent with rings and are of interest to catalysis researchers even though they are not classically considered rings. One example is the 6-membered cycle referred to as the \(\alpha\)-6-MR in literature (\cref{fig:mfi-6}) and is present in a number of frameworks including but not limited to MOR, FER, MFI, and BEA \cite{dedecek-siting-2012,bernauer-proton-2016}, which is a potential location for Co\(^{\text{2+}}\) uptake when two Al atoms are 3rd nearest neighbor (NN) in the cycle  \cite{bernauer-proton-2016,nimlos-experimental-2020}. Similar to Co\(^{\text{2+}}\) uptake at 3NN Al atoms in 6-MRs in other frameworks such as CHA \cite{di-iorio-cooperative-2020}. \cref{fig:mfi-6} shows that this particular structure would be considered two 5-MRs using connectivity rules based on a shortcut.

@@

\subsection{Characterizing T-sites by Rings}
\label{sec:orge809831}

@@

Take for example the AFI framework, containing one symmetry distinct T-site. AFI contains 4-, 6-, 12-, and 14-MRs. To describe that T-site we can count how many of each of those rings pass through the T-site. We can also prune this list using our modified definition of a shortcut, the shortest path rings definition \cite{sastre-topological-2009}, or the rings contained within the vertex symbol of this T-site \cite{okeeffe-vertex-1997}. Using the ring index outlined above, each of these conventions will provide a different description of the zeolite (highlighted in \cref{fig:afi-funnel}:
\begin{itemize}
\item Rings: 4\(\cdot\)6\(_{\text{13}} \cdot\)12\(\cdot\)14\(_{\text{7}}\)
\item This work: 4\(\cdot\)6\(_{\text{13}} \cdot\)12
\item Shortest Path Rings: 4\(\cdot\)6\(_{\text{13}}\)
\item Vertex Symbol Rings: 4\(\cdot\)6\(_{\text{11}}\)
\end{itemize}

@@

\begin{table}
\centering
\begin{threeparttable}
\caption{Comparison of ring indices for the T-sites in various uninodal zeolite frameworks. \label{tab:uninodal}}
{\scriptsize
\begin{tabular}{lllll}
\hline
Framework & Rings & This Work & Shortest Path Rings \cite{sastre-topological-2009} & Vertex Symbol Rings \cite{baerlocher-database-nodate}\\
\hline
ABW & 4\(_{\text{2}} \cdot\)6\(_{\text{3}} \cdot\)8\(_{\text{4}}\) & 4\(_{\text{2}} \cdot\)6\(_{\text{3}} \cdot\)8\(_{\text{4}}\) & 4\(_{\text{2}} \cdot\)6\(_{\text{3}} \cdot\)8\(_{\text{4}}\) & 4\(_{\text{2}} \cdot\)6\(_{\text{3}} \cdot\)8\(_{\text{2}}\)\\
ACO & 4\(_{\text{3}} \cdot\)6\(_{\text{3}} \cdot\)8\(_{\text{6}} \cdot\)10\(_{\text{15}}\) & 4\(_{\text{3}} \cdot\)8\(_{\text{6}}\) & 4\(_{\text{3}} \cdot\)8\(_{\text{6}}\) & 4\(_{\text{3}} \cdot\)8\(_{\text{6}}\)\\
AFI & 4\(_{\text{1}} \cdot\)\(_{\text{13}} \cdot\)12\(_{\text{1}} \cdot\)14\(_{\text{7}}\) & 4\(_{\text{1}} \cdot\)6\(_{\text{13}} \cdot\)12\(_{\text{1}}\) & 4\(_{\text{1}} \cdot\)6\(_{\text{13}}\) & 4\(_{\text{1}} \cdot\)6\(_{\text{11}}\)\\
ANA & 4\(_{\text{2}} \cdot\)6\(_{\text{2}} \cdot\)8\(_{\text{16}}\) & 4\(_{\text{2}} \cdot\)6\(_{\text{2}} \cdot\)8\(_{\text{16}}\) & 4\(_{\text{2}} \cdot\)6\(_{\text{2}} \cdot\)8\(_{\text{16}}\) & 4\(_{\text{2}} \cdot\)6\(_{\text{2}} \cdot\)8\(_{\text{8}}\)\\
ATO & 4\(_{\text{1}} \cdot\)6\(_{\text{9}} \cdot\)8\(_{\text{8}} \cdot\)12\(_{\text{20}}\) & 4\(_{\text{1}} \cdot\)6\(_{\text{9}} \cdot\)12\(_{\text{20}}\) & 4\(_{\text{1}} \cdot\)6\(_{\text{9}}\) & 4\(_{\text{1}} \cdot\)6\(_{\text{9}}\)\\
BCT & 4\(_{\text{1}} \cdot\)6\(_{\text{6}} \cdot\)8\(_{\text{20}}\) & 4\(_{\text{1}} \cdot\)6\(_{\text{6}} \cdot\)8\(_{\text{12}}\) & 4\(_{\text{1}} \cdot\)6\(_{\text{6}}\) & 4\(_{\text{1}} \cdot\)6\(_{\text{6}}\)\\
CHA & 4\(_{\text{3}} \cdot\)6\(_{\text{1}} \cdot\)8\(_{\text{6}} \cdot\)12\(_{\text{1}}\) & 4\(_{\text{3}} \cdot\)6\(_{\text{1}} \cdot\)8\(_{\text{2}} \cdot\)12\(_{\text{1}}\) & 4\(_{\text{3}} \cdot\)6\(_{\text{1}} \cdot\)8\(_{\text{2}}\) & 4\(_{\text{3}} \cdot\)6\(_{\text{1}} \cdot\)8\(_{\text{2}}\)\\
DFT & 4\(_{\text{2}} \cdot\)6\(_{\text{6}} \cdot\)8\(_{\text{10}} \cdot\)10\(_{\text{10}}\) & 4\(_{\text{2}} \cdot\)6\(_{\text{6}} \cdot\)8\(_{\text{10}}\) & 4\(_{\text{2}} \cdot\)6\(_{\text{6}} \cdot\)8\(_{\text{10}}\) & 4\(_{\text{2}} \cdot\)6\(_{\text{4}} \cdot\)8\(_{\text{6}}\)\\
GIS & 4\(_{\text{3}} \cdot\)8\(_{\text{4}}\) & 4\(_{\text{3}} \cdot\)8\(_{\text{4}}\) & 4\(_{\text{3}} \cdot\)8\(_{\text{4}}\) & 4\(_{\text{3}} \cdot\)8\(_{\text{4}}\)\\
GME & 4\(_{\text{3}} \cdot\)6\(_{\text{1}} \cdot\)8\(_{\text{6}} \cdot\)12\(_{\text{7}}\) & 4\(_{\text{3}} \cdot\)6\(_{\text{1}} \cdot\)8\(_{\text{2}} \cdot\)12\(_{\text{1}}\) & 4\(_{\text{3}} \cdot\)6\(_{\text{1}} \cdot\)8\(_{\text{2}}\) & 4\(_{\text{3}} \cdot\)6\(_{\text{1}} \cdot\)8\(_{\text{2}}\)\\
MER & 4\(_{\text{3}} \cdot\)8\(_{\text{4}} \cdot\)10\(_{\text{10}} \cdot\)14\(_{\text{14}}\) & 4\(_{\text{3}} \cdot\)8\(_{\text{4}}\) & 4\(_{\text{3}} \cdot\)8\(_{\text{4}}\) & 4\(_{\text{3}} \cdot\)8\(_{\text{4}}\)\\
MON & 4\(_{\text{1}} \cdot\)5\(_{\text{5}} \cdot\)8\(_{\text{6}}\) & 4\(_{\text{1}} \cdot\)5\(_{\text{5}} \cdot\)8\(_{\text{6}}\) & 4\(_{\text{1}} \cdot\)5\(_{\text{5}} \cdot\)8\(_{\text{6}}\) & 4\(_{\text{1}} \cdot\)5\(_{\text{4}} \cdot\)8\(_{\text{4}}\)\\
NPO & 3\(_{\text{1}} \cdot\)6\(_{\text{6}} \cdot\)12\(_{\text{40}}\) & 3\(_{\text{1}} \cdot\)6\(_{\text{6}} \cdot\)12\(_{\text{40}}\) & 3\(_{\text{1}} \cdot\)6\(_{\text{6}}\) & 3\(_{\text{1}} \cdot\)6\(_{\text{6}}\)\\
\hline
\end{tabular}
\begin{tablenotes}
Vertex symbols have been represented in ring index format for ease of comparison.
}
\end{tablenotes}
\end{threeparttable}
\end{table}

@@

To complete the comparison of ring counting conventions, we have developed method to determine how similar the ring counts of one convention are to another. We do this by comparing the ring index for a T-site using each convention, where the similarity of the ring indices is scored with \cref{eq:similarity}. In this equation, sr is the number of similar rings that are found in both counting conventions, and mr is the maximum number of rings found by either convention. For example: the ring index of AFI using the classic shortcut definition and the shortest path definition are 4\(\cdot\)6\(_{\text{13}} \cdot\)12\(\cdot\)14\(_{\text{7}}\) and 4\(\cdot\)6\(_{\text{13}} \cdot\). The number of similar rings found by both conventions is 14, and the maximum number of rings found by either convention is 22. This would lead to a similarity score of 0.636. We do this for every T-site between two conventions and average the similarity score to get the results in \cref{fig:similarity}. Down the diagonal each method is compared to itself and clearly has a similarity of 1. The remainder of the table follows intuition, in that the  most restrictive ring counting convention (vertex symbol rings) compare to the least restrictive convention (rings) has the lowest similarity score. The two most similar ring counting methods are our modified shortcut definition and the shortest path rings.

@@

One final point we would like to make about T-site characterization, is that the vertex symbol and ring indices only provide information about rings, and do not give any information about the spatial orientation of those rings around the T-site. The vertex symbol comes close to accomplishing that, by finding rings associated with opposite edge pairs of the tetrahedron, and then listing the those pairs from smallest to largest. However, the vertex symbol does not capture subtle but distinct differences in the orientation of the rings around the T-site that can lead to varying local void environments. For example: MOR T3, MON T1, and EON T9 all have the same vertex symbol of: 4\codt5\(_{\text{2}} \cdot\)5\(\cdot\)8\(_{\text{2}} \cdot\)5\(\cdot\)8\(_{\text{2}}\). However, the orientation of those rings around each of those T-sites are not identical. \cref{fig:stereo} shows a cutout of each of these frameworks that only includes the atoms that make up the rings of the vertex symbol around the specified T-sites. We can see that MOR T3 and EON T9 have the same ring orientation, and that orientation is different from the rings making up the vertex symbol of MON T1. The main difference is the location of the 5\(_{\text{2}}\)- and 4-MR edges. They are highlighted in \cref{fig:stereo}.

@@

With this new descriptor MOR T3 and EON T9 would be labeled as: 8\(_{\text{2}} \cdot\)8\(_{\text{2}} \cdot\)5\(_{\text{2}} \cdot\)5\(\cdot\)4\(\cdot\)5, and MON T1 as: 8\(_{\text{2}} \cdot\)8\(_{\text{2}} \cdot\)4\(\cdot\)5\(\cdot\)5\(_{\text{2}} \cdot\)5. The difference is subtle but highlights the distinct structural difference between the two types of T-sites that is not otherwise captured by a vertex symbol. Stereochemistry of the rings associated with a T-site could influence the chemical properties we care about such as deprotonation energy, T-site substitution energy, or catalytic properties of reactions happening at that T-site.

@@

\section{Acknowledgments}
\label{sec:org805a78b}

>>>>

% Created 2022-10-03 Mon 12:32
% Intended LaTeX compiler: pdflatex
\documentclass[preprint,numrefs,noinfo,sort&compress]{elsarticle}
  \usepackage[utf8]{inputenc}
\usepackage{url}
\usepackage[version=4]{mhchem}
\usepackage{chemmacros}[2016/05/02]
\usepackage{graphicx}
\usepackage{float}
\usepackage{color}
\usepackage{adjustbox}
\usepackage{amsmath}
\usepackage{siunitx}
\usepackage{textcomp}
\usepackage{wasysym}
\usepackage{latexsym}
\usepackage{amssymb}
\usepackage{lineno}
\usepackage{chemformula}
\usepackage{xr}
\usepackage{pifont}
\usepackage{longtable}
\usepackage[section]{placeins}
\usepackage{threeparttable}
\newcommand{\red}[1]{\textcolor{red}{#1}}
\chemsetup{formula = mhchem ,modules = {reactions,thermodynamics}}
\usepackage{xcolor}
\chemsetup[reactions]{tag-open= ( , tag-close = )}
\usepackage[capitalize]{cleveref}
\usepackage{minted}
\DeclareSIUnit\Td{Td}
\setminted[python]{frame=lines,fontsize=\scriptsize,xleftmargin=\parindent,linenos,breaklines}
\date{}
\title{}
\begin{document}

@@

\mainmatter
\section{Introduction}
\label{sec:org7af04cb}
Zeolites are three dimensional crystalline structures containing tetrahedral Si or Al atoms connected by oxygen bridges. The International Zeolite Association (IZA) lists over 200 known zeolite frameworks that are described by dimensionality, pore shape and size, and Si/Al ratios \cite{baerlocher-database-nodate}. It is natural to characterize differences in zeolites by the size and shapes of the features present in the crystal. Rings are one common type of feature widely reported and used, both to characterize a zeolite crystal and as descriptors of the individual tetrahedral sites (T-sites) of a zeolite. Rings have been used to identify feasible zeolites topologies \cite{li-why-2014}, to describe the similarity and differences between zeolites \cite{curtis-statistical-2003,bermudez-calculation-2017}, to identify sites or voids of catalytic relevance \cite{li-first-principles-2018,kester-effects-2021}, and as machine learning finger prints of physicochemical properties \cite{helfrecht-new-2019}.

@@

Vertex symbols are a common way to describe the T-sites of a zeolite framework, and were first used to do so in 1997 by O'Keeffe and Hyde in 1997 \cite{okeeffe-vertex-1997}. The vertex symbol of a T-site contains the shortest rings associated with each of the six edges of the tetrahedron formed by the T-atom and its four bounded oxygen atoms. The ring sizes, and  their multiplicity, for opposite edges (edges of the tetrahedron that do not connect to the same oxygen atom) are grouped together. These grouped pairs are listed from smallest to largest forming the vertex symbol. \cref{fig:cha-vertex}(a) shows an example of the tetrahedron formed at the single symmetry distinct T-site in the CHA framework. The edges of the tetrahedron are labeled to aid in identification. \cref{fig:cha-vertex}(b-d) show the rings associated with each opposite pair of edges. The vertex symbol of T1 in CHA is thus able to be determined as 4\(\cdot\)4\(\cdot\)4\(\cdot\)8\(\cdot\)6\(\cdot\)8. For a T-site that contains a multiplicity of rings at one edge, that multiplicity would be represented as a subscript in the vertex symbol. An example would be the vertex symbol of 4\(\cdot\)6\textsubscript{2}\(\cdot\)6\(\cdot\)6\textsubscript{3}\(\cdot\)6\textsubscript{2}\(\cdot\)6\textsubscript{3} for T1 in AFI.

@@

\section{Methods}
\label{sec:org83ee037}

\subsection{Finding Rings That do not Contain Shortcuts \label{section:goetzke}}
\label{sec:orge9b7691}

@@

\subsection{Finding Vertex Symbol Rings \label{section:vertex}}
\label{sec:org89c7798}

@@

\subsection{Finding Shortest Path Rings}
\label{sec:org0530c72}

@@

\subsection{Modified Shortcut Rings}
\label{sec:org7cfe2eb}

@@

\subsection{Ordered Vertex Symbols \label{section:ov}}
\label{sec:org54f88e7}

@@

\subsection{Determining All Ring Sizes Contained in a Zeolite Framework}
\label{sec:orgb8842aa}

@@

\section{Results}
\label{sec:org3113701}
\subsection{Characterizing Rings in a Zeolite Graph}
\label{sec:org11cd53a}
The IZA Database \cite{baerlocher-database-nodate} is a common reference used to identify all the rings in a zeolite framework, however it only lists the rings that define a channel (ex: 12-MR in AFI), or rings associated with the symbol of a T-site. These rings listed by the IZA are referred to as tabulated rings in the literature \cite{curtis-statistical-2003}. In some frameworks, other rings (cycles not containing shortcuts) exist that are not included in the list of tabulated rings. These 'untabulated' rings may still provide important topological information about a zeolite framework, or the local void environment around a T-site. \cref{fig:fw-counts} shows counts of frameworks containing each size ring from 3- to 18-MR using the Goetzke algorithm and the listed rings on the IZA database \cite{baerlocher-database-nodate}. There are slight differences in the counts up to 6-MRs, but the main divergence takes place as we get to ring sizes \textgreater6-MR.

@@

\subsection{Characterizing Frameworks by Rings}
\label{sec:orgc427fb9}
With the addition of our modified shortcut definition, we have four ring finding conventions to compare, as well as including the tabulated rings from the IZA Database. \cref{fig:ring-counts} shows how many frameworks contain each size ring found using the various ring counting conventions from 3- to 18-MRs. This plot highlights the differences in the conventions and shows that a topological description of a framework based on rings will depend on the way that you define a ring. In general, a hierarchy of ring sizes found by each convention is: all rings not containing a shortcut \textgreater this work \textgreater shortest path rings \textgreater vertex symbol rings. While the IZA listed rings includes the vertex symbol rings, and a selection of general rings \cite{baerlocher-database-nodate}.

@@

One drawback to using a ring convention based on connectivity and shortcuts is the exclusion of cycles that don't fit this definition, but still exhibit properties like rings. This is a trade-off between well-defined connectivity rules, and the inclusion of particular void environments that may still have important applications. These shortcut containing cycles can display chemical and/or geometric properties consistent with rings and are of interest to catalysis researchers even though they are not classically considered rings. One example is the 6-membered cycle referred to as the \(\alpha\)-6-MR in literature (\cref{fig:mfi-6}) and is present in a number of frameworks including but not limited to MOR, FER, MFI, and BEA \cite{dedecek-siting-2012,bernauer-proton-2016}, which is a potential location for Co\textsuperscript{2+} uptake when two Al atoms are 3rd nearest neighbor (NN) in the cycle  \cite{bernauer-proton-2016,nimlos-experimental-2020}. Similar to Co\textsuperscript{2+} uptake at 3NN Al atoms in 6-MRs in other frameworks such as CHA \cite{di-iorio-cooperative-2020}. \cref{fig:mfi-6} shows that this particular structure would be considered two 5-MRs using connectivity rules based on a shortcut.

@@

\subsection{Characterizing T-sites by Rings}
\label{sec:orgc94a591}

@@

Take for example the AFI framework, containing one symmetry distinct T-site. AFI contains 4-, 6-, 12-, and 14-MRs. To describe that T-site we can count how many of each of those rings pass through the T-site. We can also prune this list using our modified definition of a shortcut, the shortest path rings definition \cite{sastre-topological-2009}, or the rings contained within the vertex symbol of this T-site \cite{okeeffe-vertex-1997}. Using the ring index outlined above, each of these conventions will provide a different description of the zeolite (highlighted in \cref{fig:afi-funnel}:
\begin{itemize}
\item Rings: 4\(\cdot\)6\textsubscript{13}\(\cdot\)12\(\cdot\)14\textsubscript{7}
\item This work: 4\(\cdot\)6\textsubscript{13}\(\cdot\)12
\item Shortest Path Rings: 4\(\cdot\)6\textsubscript{13}
\item Vertex Symbol Rings: 4\(\cdot\)6\textsubscript{11}
\end{itemize}

@@

\begin{table}
\centering
\begin{threeparttable}
\caption{Comparison of ring indices for the T-sites in various uninodal zeolite frameworks. \label{tab:uninodal}}
{\scriptsize
\begin{tabular}{lllll}
\hline
Framework & Rings & This Work & Shortest Path Rings \cite{sastre-topological-2009} & Vertex Symbol Rings \cite{baerlocher-database-nodate}\\
\hline
ABW & 4\textsubscript{2}\(\cdot\)6\textsubscript{3}\(\cdot\)8\textsubscript{4} & 4\textsubscript{2}\(\cdot\)6\textsubscript{3}\(\cdot\)8\textsubscript{4} & 4\textsubscript{2}\(\cdot\)6\textsubscript{3}\(\cdot\)8\textsubscript{4} & 4\textsubscript{2}\(\cdot\)6\textsubscript{3}\(\cdot\)8\textsubscript{2}\\
ACO & 4\textsubscript{3}\(\cdot\)6\textsubscript{3}\(\cdot\)8\textsubscript{6}\(\cdot\)10\textsubscript{15} & 4\textsubscript{3}\(\cdot\)8\textsubscript{6} & 4\textsubscript{3}\(\cdot\)8\textsubscript{6} & 4\textsubscript{3}\(\cdot\)8\textsubscript{6}\\
AFI & 4\textsubscript{1}\(\cdot\)\textsubscript{13}\(\cdot\)12\textsubscript{1}\(\cdot\)14\textsubscript{7} & 4\textsubscript{1}\(\cdot\)6\textsubscript{13}\(\cdot\)12\textsubscript{1} & 4\textsubscript{1}\(\cdot\)6\textsubscript{13} & 4\textsubscript{1}\(\cdot\)6\textsubscript{11}\\
ANA & 4\textsubscript{2}\(\cdot\)6\textsubscript{2}\(\cdot\)8\textsubscript{16} & 4\textsubscript{2}\(\cdot\)6\textsubscript{2}\(\cdot\)8\textsubscript{16} & 4\textsubscript{2}\(\cdot\)6\textsubscript{2}\(\cdot\)8\textsubscript{16} & 4\textsubscript{2}\(\cdot\)6\textsubscript{2}\(\cdot\)8\textsubscript{8}\\
ATO & 4\textsubscript{1}\(\cdot\)6\textsubscript{9}\(\cdot\)8\textsubscript{8}\(\cdot\)12\textsubscript{20} & 4\textsubscript{1}\(\cdot\)6\textsubscript{9}\(\cdot\)12\textsubscript{20} & 4\textsubscript{1}\(\cdot\)6\textsubscript{9} & 4\textsubscript{1}\(\cdot\)6\textsubscript{9}\\
BCT & 4\textsubscript{1}\(\cdot\)6\textsubscript{6}\(\cdot\)8\textsubscript{20} & 4\textsubscript{1}\(\cdot\)6\textsubscript{6}\(\cdot\)8\textsubscript{12} & 4\textsubscript{1}\(\cdot\)6\textsubscript{6} & 4\textsubscript{1}\(\cdot\)6\textsubscript{6}\\
CHA & 4\textsubscript{3}\(\cdot\)6\textsubscript{1}\(\cdot\)8\textsubscript{6}\(\cdot\)12\textsubscript{1} & 4\textsubscript{3}\(\cdot\)6\textsubscript{1}\(\cdot\)8\textsubscript{2}\(\cdot\)12\textsubscript{1} & 4\textsubscript{3}\(\cdot\)6\textsubscript{1}\(\cdot\)8\textsubscript{2} & 4\textsubscript{3}\(\cdot\)6\textsubscript{1}\(\cdot\)8\textsubscript{2}\\
DFT & 4\textsubscript{2}\(\cdot\)6\textsubscript{6}\(\cdot\)8\textsubscript{10}\(\cdot\)10\textsubscript{10} & 4\textsubscript{2}\(\cdot\)6\textsubscript{6}\(\cdot\)8\textsubscript{10} & 4\textsubscript{2}\(\cdot\)6\textsubscript{6}\(\cdot\)8\textsubscript{10} & 4\textsubscript{2}\(\cdot\)6\textsubscript{4}\(\cdot\)8\textsubscript{6}\\
GIS & 4\textsubscript{3}\(\cdot\)8\textsubscript{4} & 4\textsubscript{3}\(\cdot\)8\textsubscript{4} & 4\textsubscript{3}\(\cdot\)8\textsubscript{4} & 4\textsubscript{3}\(\cdot\)8\textsubscript{4}\\
GME & 4\textsubscript{3}\(\cdot\)6\textsubscript{1}\(\cdot\)8\textsubscript{6}\(\cdot\)12\textsubscript{7} & 4\textsubscript{3}\(\cdot\)6\textsubscript{1}\(\cdot\)8\textsubscript{2}\(\cdot\)12\textsubscript{1} & 4\textsubscript{3}\(\cdot\)6\textsubscript{1}\(\cdot\)8\textsubscript{2} & 4\textsubscript{3}\(\cdot\)6\textsubscript{1}\(\cdot\)8\textsubscript{2}\\
MER & 4\textsubscript{3}\(\cdot\)8\textsubscript{4}\(\cdot\)10\textsubscript{10}\(\cdot\)14\textsubscript{14} & 4\textsubscript{3}\(\cdot\)8\textsubscript{4} & 4\textsubscript{3}\(\cdot\)8\textsubscript{4} & 4\textsubscript{3}\(\cdot\)8\textsubscript{4}\\
MON & 4\textsubscript{1}\(\cdot\)5\textsubscript{5}\(\cdot\)8\textsubscript{6} & 4\textsubscript{1}\(\cdot\)5\textsubscript{5}\(\cdot\)8\textsubscript{6} & 4\textsubscript{1}\(\cdot\)5\textsubscript{5}\(\cdot\)8\textsubscript{6} & 4\textsubscript{1}\(\cdot\)5\textsubscript{4}\(\cdot\)8\textsubscript{4}\\
NPO & 3\textsubscript{1}\(\cdot\)6\textsubscript{6}\(\cdot\)12\textsubscript{40} & 3\textsubscript{1}\(\cdot\)6\textsubscript{6}\(\cdot\)12\textsubscript{40} & 3\textsubscript{1}\(\cdot\)6\textsubscript{6} & 3\textsubscript{1}\(\cdot\)6\textsubscript{6}\\
\hline
\end{tabular}
\begin{tablenotes}
Vertex symbols have been represented in ring index format for ease of comparison.
}
\end{tablenotes}
\end{threeparttable}
\end{table}

@@

To complete the comparison of ring counting conventions, we have developed method to determine how similar the ring counts of one convention are to another. We do this by comparing the ring index for a T-site using each convention, where the similarity of the ring indices is scored with \cref{eq:similarity}. In this equation, sr is the number of similar rings that are found in both counting conventions, and mr is the maximum number of rings found by either convention. For example: the ring index of AFI using the classic shortcut definition and the shortest path definition are 4\(\cdot\)6\textsubscript{13}\(\cdot\)12\(\cdot\)14\textsubscript{7} and 4\(\cdot\)6\textsubscript{13}\(\cdot\). The number of similar rings found by both conventions is 14, and the maximum number of rings found by either convention is 22. This would lead to a similarity score of 0.636. We do this for every T-site between two conventions and average the similarity score to get the results in \cref{fig:similarity}. Down the diagonal each method is compared to itself and clearly has a similarity of 1. The remainder of the table follows intuition, in that the  most restrictive ring counting convention (vertex symbol rings) compare to the least restrictive convention (rings) has the lowest similarity score. The two most similar ring counting methods are our modified shortcut definition and the shortest path rings.

@@

One final point we would like to make about T-site characterization, is that the vertex symbol and ring indices only provide information about rings, and do not give any information about the spatial orientation of those rings around the T-site. The vertex symbol comes close to accomplishing that, by finding rings associated with opposite edge pairs of the tetrahedron, and then listing the those pairs from smallest to largest. However, the vertex symbol does not capture subtle but distinct differences in the orientation of the rings around the T-site that can lead to varying local void environments. For example: MOR T3, MON T1, and EON T9 all have the same vertex symbol of: 4\codt5\textsubscript{2}\(\cdot\)5\(\cdot\)8\textsubscript{2}\(\cdot\)5\(\cdot\)8\textsubscript{2}. However, the orientation of those rings around each of those T-sites are not identical. \cref{fig:stereo} shows a cutout of each of these frameworks that only includes the atoms that make up the rings of the vertex symbol around the specified T-sites. We can see that MOR T3 and EON T9 have the same ring orientation, and that orientation is different from the rings making up the vertex symbol of MON T1. The main difference is the location of the 5\textsubscript{2}- and 4-MR edges. They are highlighted in \cref{fig:stereo}.

@@

With this new descriptor MOR T3 and EON T9 would be labeled as: 8\textsubscript{2}\(\cdot\)8\textsubscript{2}\(\cdot\)5\textsubscript{2}\(\cdot\)5\(\cdot\)4\(\cdot\)5, and MON T1 as: 8\textsubscript{2}\(\cdot\)8\textsubscript{2}\(\cdot\)4\(\cdot\)5\(\cdot\)5\textsubscript{2}\(\cdot\)5. The difference is subtle but highlights the distinct structural difference between the two types of T-sites that is not otherwise captured by a vertex symbol. Stereochemistry of the rings associated with a T-site could influence the chemical properties we care about such as deprotonation energy, T-site substitution energy, or catalytic properties of reactions happening at that T-site.

@@

\section{Acknowledgments}
\label{sec:org5f43fe4}
